---
title: "PA #4: Graph Revision"
author: "Andrew Nalundasan"
date: "7/21/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Course: OMSBA 5210, Seattle University
# Purpose: Participation Assignment 4
# Date: July 21, 2021
# Author: Andrew Nalundasan
```

```{r, include=FALSE}
# The tidyverse package contains ggplot2, dplyr, and several other packages
library(tidyverse)

# to get $ and , on axis in graph
library(scales)

# NHK special
library(vtable)

# to make annotations
library(ggannotate)
```


```{r, include=FALSE}
# Load and prepare data 
zillow <- read_csv("../02_raw_data/qp1_data.csv")

# look at the data
head(zillow)
vtable(zillow)
```


# Revised Visualization

Revision of waterfront scatterplot (3rd graph out of 4 in assignment)

+ Changed color of waterfront to blue since blue is associated with water
+ Set transparency to 0.35 to see overlapping data points
+ Removed tick marks to reduce clutter
+ Removed border to reduce clutter
+ Changed y axis to show price in millions and removed trailing 0's
+ Reworded title and subtitle

```{r}
cb_palette <- c("#999999", "#00A5FF")

p <- zillow %>% 
  group_by(sqft_living, waterfront) %>% 
  summarize(med_price = median(price)) %>% 
  filter(sqft_living < 9000) %>% 
  mutate(waterfront = as.factor(waterfront)) %>% 
  
  ggplot(mapping = aes(x = sqft_living, y = med_price, color = waterfront)) + 
  scale_color_manual(values = cb_palette) + 
  geom_point(alpha = 0.35) + 
  geom_smooth(method = "lm", se = FALSE) + 
  
  # scale_y_continuous(labels = dollar) + 
  scale_y_continuous(label = label_number_si(accuracy = .5)) +
  guides(color = "none", labels = "none") + 
  theme_classic() + 
  #theme(axis.line.x = element_blank()) + 
  #theme(axis.line.y = element_blank()) + 
  theme(axis.ticks.x = element_blank()) + 
  theme(axis.ticks.y = element_blank()) +
  
  labs(x = "Living Space (in Square Feet)",
       y = "Median Price",
       title = "Waterfront property boosts home price", 
       subtitle = "More living space paired with waterfront views\ncompound value of home") + 
  
  geom_text(data = data.frame(x = 7840.10833134011, y = 4134345.8884987, label = "Waterfront"),
            mapping = aes(x = x, y = y, label = label),
            hjust = 0L, vjust = 0.35, colour = "#00A5FF", fontface = 2, inherit.aes = FALSE) +
  
  geom_text(data = data.frame(x = 8019.8032352691, y = 2169630.60683532, label = "No Waterfront"),
            mapping = aes(x = x, y = y, label = label),
            hjust = 0.32, vjust = 1L, colour = "#999999", fontface = 2, inherit.aes = FALSE)


# view the viz
p

# make annotations
# ggannotate(p)

# save the viz
ggsave(filename = "PA4.png", plot = p)
```

