---
title: "PA5"
author: "Andrew Nalundasan"
date: "7/21/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
# load libraries
library(tidyverse)
library(scales)
library(janitor)

```


```{r, include=FALSE}
# Load and prepare data 
zillow <- read_csv("../02_raw_data/qp1_data.csv")
```

# Steps: 

1. statistically test the main two findings you presented in your analysis

2. state whether each finding is:

    - practically significant
    - statistically significant

3. write a technical sentence describing the statistical outcome for each finding

4. write a non-technical sentence describing the statistical outcome for each finding 
       
    - make sure it is appropriate for a non-technical audience

5. copy and paste the results and sentences into the assignment page on Canvas

# Main finding 1: the King County housing market is flooded by grade 7, grade 8, and grade 9 houses

```{r}
# proportion of counts
zillow %>% 
  filter(yr_renovated > 1975) %>% 
  mutate(grade = as.factor(grade)) %>%
  ggplot(mapping = aes(x = yr_renovated, fill = grade)) + 
  geom_bar(position = "fill") + 
  labs(x = "Year Renovated", 
       y = "Proportion", 
       title = "Proportion or grade rating with year renovated")



```

```{r}
# t-test of grade on year renovated
# create new object to filter yr_renovated to be everything post 1975
post_1975 <- zillow %>% 
  filter(yr_renovated > 1975)

# t-test
(t <- t.test(post_1975$yr_renovated, post_1975$grade, conf.level = 0.95))
```

**Comments** 

+ Ho: grade of home is the same for each year

+ Practical Significance:

    - mean grade of home is 7.83, so housing market has home grades > 7
    
+ Statistical Significance:

    - p-value is really small, so this is consistent
    
+ Technical sentence:

    - We reject the null that home grade is the same from year to year

+ Non-technical sentence:

    - Grade of home changes from year to year

# Main finding 2: bathrooms and grade greatly impact housing prices

```{r}
# price relationship differences

bath <-  c("1", "1.5", "2", "2.5", "3", "3.5", "4")
bath <- as.factor(bath)

zillow %>% 
  filter(price < 4000000) %>% 
  filter(bathrooms %in% bath) %>% 
  mutate(bathrooms = as.factor(bathrooms)) %>% 
  ggplot(mapping = aes(x = grade, y = price)) +
  geom_point(position = "jitter") +
  scale_y_continuous(labels = dollar) + 
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) + 
  geom_smooth(mapping = aes(color = bathrooms), method = "lm", se = FALSE) + 
  labs(x = "Grade", 
       y = "Price", 
       title = "Home pricing trends with respect to Grades and Bathrooms")
```

```{r}
# Number of obs helps us determine if stats will be valid
bath <-  c("1", "1.5", "2", "2.5", "3", "3.5", "4")
bath <- as.factor(bath)

zillow_bath <- zillow %>% 
  filter(bathrooms %in% bath)

zillow_bath %>% 
  tabyl(grade, bathrooms)
```

**Comments**

+ Grades 7, 8, and 9 have highest counts of all bathroom categories

+ H0: Home grades are the same by number of bathrooms

```{r}
# Using chi-square tests for counts we can only determine if they are the same or not
# Are grades similar by number of bathrooms?
# Not tidyverse so have to use base code

chisq.test(table(zillow_bath$grade, zillow_bath$bathrooms))
```
**Comments**

+ Practical Significance:

    - Is this practically significant?
    
+ Statistical Significance:

    - chi-square: large, so reject the null
    - p-value: small, so reject the null
    
+ Technical sentence:

    - We reject the null hypothesis that home grades are the same regardless of number of bathrooms a home has. 

+ Non-technical sentence:

    - Home grades differ depending on the number of bathrooms that a home has.
